Backpropagating Gravitational wave events

### 2022-04-16 src/ms.tex
update

Once all the parameters including a particular draw of all the random variables
represented by $\bm{X}$ are known, the population synthesis code is a
deterministic function that transforms the properties of the progenitors into
the GW parameters $\bm{\theta}$:
\begin{equation}
    \bm{\theta} = \bm{f}\left( \bm{\Theta} \right).
\end{equation}
The mapping $\bm{f}$ from $\bm{\Theta}$ to $\bm{\theta}$ is many-to-one.  
In order to draw inferences about $\bm{\Theta}$ from gravitational wave data, we
need to be able to evaluate the likelihood of that data at fixed progenitor
parameters, 
\begin{equation}
    p\left( d \mid \bm{\Theta} \right) = p\left( d \mid \bm{\theta} = \bm{f}\left( \bm{\Theta} \right) \right),
\end{equation}
because the likelihood depends only on the gravitational waveform generated by
parameters $\bm{\theta}$ to which the progenitor parameters map.  In principle
this likelihood could be computed at arbitrary vaues of $\bm{\Theta}$ using the
same machinery that is used to estimate source parameters in gravitational wave
catalogs \citep{Veitch2015,Ashton2019,Romero-Shaw2020,GWTC-3} to develop samples
of progenitor parameters, hyper-parameters, and random variables in
$\bm{\Theta}$ from a posterior density.

However, it is very expensive to couple progenitor evolution to stochastic
sampling in this way.  Because the gravitational wave likelihood function
ultimately depends only on the GW parameters $\theta$, and catalogs have already
produced samples over these parameters from a posterior density 
\begin{equation}
    p\left( \bm{\theta} \mid d \right) \propto p\left( d \mid \bm{\theta} \right) p_\mathrm{GW} \left( \bm{\theta} \right),
\end{equation}
where $p_\mathrm{GW}$ is the prior density used for sampling, it is much more
efficient to \wf{WIP}.

>>>>

Once all the parameters including a particular draw of all the random variables
represented by $\bm{X}$ are known, the population synthesis code is a
deterministic function that transforms the properties of the progenitors into
the GW parameters $\bm{\theta}$:
\begin{equation}
    \bm{\theta} = \bm{f}\left( \bm{\Theta} \right).
\end{equation}
The mapping $\bm{f}$ from $\bm{\Theta}$ to $\bm{\theta}$ can be many-to-one.  
In order to draw inferences about $\bm{\Theta}$ from gravitational wave data, we
need to be able to evaluate the likelihood of that data at fixed progenitor
parameters, namely
\begin{equation}
    p\left( d \mid \bm{\Theta} \right) = p\left( d \mid \bm{\theta} = \bm{f}\left( \bm{\Theta} \right) \right).
\end{equation}
The equality holds because the likelihood depends only on the gravitational waveform generated by
parameters $\bm{\theta}$ to which the progenitor parameters map onto.  In principle
this likelihood could be computed at arbitrary values of $\bm{\Theta}$ using the
same machinery that is used to estimate source parameters in gravitational wave
catalogs \citep{Veitch2015,Ashton2019,Romero-Shaw2020,GWTC-3} to develop samples
of progenitor parameters, hyper-parameters, and random variables in
$\bm{\Theta}$ from a posterior density.

\kw{Talk about the MCMC}

Since the gravitational wave likelihood function
ultimately depends only on the GW parameters $\theta$, and GWTC3 have already
produced samples over these parameters from a posterior density 
\begin{equation}
    p\left( \bm{\theta} \mid d \right) \propto p\left( d \mid \bm{\theta} \right) \pi_\mathrm{GW} \left( \bm{\theta} \right),
\end{equation}
where $\pi_\mathrm{GW}$ is the prior density used for sampling,
we can save the computation cost in evaluating the GW likelihood by rewriting the posterior density of $\Theta$ in terms of the posterior density of $\theta$, namely

\begin{align}
    p(\Theta | d) = \frac{p(\theta(\Theta)| d) \pi(\Theta)}{\pi_\mathrm{GW}(\theta)}.
\end{align}

Including hyper-parameters and random variables in the sampling drastically increase the dimensionality of the problem,
which makes the sampling process more computationally expensive and harder to converge.
To speed up the convergence, we
